Analysing the effect of land-use type on the proportion of scouts

### 2021-10-28 analysis/fit_models.Rmd
started converting other rmwrkdown files into having a nice htlm knitting

# Fitting waggle dance models to waggle dance data
## Joseph Palmer
---

This document contains code to fit the waggle dance model to each of our 20 different sites. The code here uses the wagglefit package, as well as some additional code stored in `fit_data.R` to help simplify things here.

```{r, include = FALSE}
devtools::load_all()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(family = "DejaVuSerif", size = 8)
    )
)
library(cowplot)
library(dplyr)
library(tibble)
source("analysis/fit_data.R")

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

```{r}

@@

@@

ggsave(
  plot = aic_plot,
  filename = "analysis/figures/AIC_plot.png",
  width = 90,
  height = 110,
  units = "mm",
  dpi = 300
)

@@

ggsave(
  plot = ks_plot_dist,
  filename = "analysis/figures/sites_ks.png",
  width = 86,
  height = 180,
  units = "mm",
  dpi = 300
)

ggsave(
  plot = all_sites$STU$fit,
  filename = "analysis/figures/STU.png",
  width = 90,
  height = 110,
  units = "mm",
  dpi = 300
)

ggsave(
  plot = all_sites$ZSL$fit,
  filename = "analysis/figures/ZSL.png",
  width = 90,
  height = 110,
  units = "mm",
  dpi = 300
)
```

@@

```{r}

@@

full_data_path <- "analysis/data/FullHBForagingData.csv"
data_raw <- tibble(read.csv(full_data_path))

@@

# london area
london <- st_read(
  "analysis/shapefiles/London_Ward.shp"
)

@@

ggsave(
  plot = merge_plot,
  filename = "analysis/figures/site_map.png",
  width = 90,
  height = 110,
  units = "mm",
  dpi = 300
)

@@

ggsave(
  plot = sites_model_plot,
  filename = "analysis/figures/sites_model_plot.png",
  width = 183,
  height = 190,
  units = "mm",
  dpi = 300
)
```

```{r}
# other map plots
inset <- ggplot() +
  geom_sf(
    data = worldmap,
    fill = "grey90",
    color = "#4b4949d0"
  ) +
  geom_sf(
    data = london,
    fill = "#ced0cffc",
    lwd = 0
  ) +
  coord_sf(
    xlim = c(points_area[[1]] - 0.1, points_area[[3]] + 0.1),
    ylim = c(points_area[[2]] - 0.1, points_area[[4]] + 0.1)
  ) +
  geom_point(
    data = map_data,
    aes(x = lon, y = lat, shape = model, colour = model), size = 1.5
  ) +
  annotation_north_arrow(
    location = "tr", which_north = "true",
    style = north_arrow_fancy_orienteering,
    height = unit(10, "mm"),
    width = unit(10, "mm")
  ) +
  annotation_scale(
    location = "br",
    text_cex = 3
  ) +
  scale_shape_manual(values = c(1, 2)) +
  scale_colour_manual(values = c("black", "red")) +
  theme_nothing() +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL, y = NULL) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = .5),
    text = element_text(size = 42)
  )

@@

ggsave(
  plot = merge_plot,
  filename = "analysis/figures/site_map.png",
  width = 183,
  height = 190,
  units = "mm",
  dpi = 300
)

@@

ggsave(
  plot = sites_model_plot,
  filename = "analysis/figures/results_model_plot.png",
  width = 183,
  height = 190,
  units = "mm",
  dpi = 300
)
```

>>>>

---
title: Fitting waggle dance models to waggle dance data
author: Joseph Palmer
output:
  html_notebook:
    theme: yeti
    toc: true
    toc_float: true
---

This document contains code to fit the waggle dance model to each of our 20 different sites. The code here uses the wagglefit package, as well as some additional code stored in `fit_data.R` to help simplify things here. Each site has its own section, with reused code. The final sections create map plots and summary statistic plots.

```{r, preamble, include = FALSE}
devtools::load_all()

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(ggplot2)
theme_set(
  theme_classic() +
    theme(
      text = element_text(family = "DejaVuSerif", size = 8)
    )
)
library(cowplot)
library(dplyr)
library(tibble)
source("fit_data.R")

@@

```{r, BEL, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, BFI, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, BLO, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, BUR, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, CAD, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, GIL, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, HER, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, HHS, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, HOR, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, MAK, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, MEL, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, MPA, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, ROT, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, SAU, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, SOM, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, SRA, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, STU, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, SWP, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, YAL, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, ZSL, include=FALSE, message = FALSE, warning = FALSE}

@@

```{r, make-plots, include=FALSE, message = FALSE, warning = FALSE}

@@

# save results
saveRDS(df, file = "results/site_fit_results.Rda")

@@

ggsave(
  plot = aic_plot,
  filename = "results/figures/AIC_plot.png",
  width = 90,
  height = 110,
  units = "mm",
  dpi = 300
)

@@

ggsave(
  plot = ks_plot_dist,
  filename = "results/figures/sites_ks.png",
  width = 86,
  height = 180,
  units = "mm",
  dpi = 300
)

ggsave(
  plot = all_sites$STU$fit,
  filename = "results/figures/STU.png",
  width = 90,
  height = 110,
  units = "mm",
  dpi = 300
)

ggsave(
  plot = all_sites$ZSL$fit,
  filename = "results/figures/ZSL.png",
  width = 90,
  height = 110,
  units = "mm",
  dpi = 300
)
```

@@

```{r, map-plots, include=FALSE, message = FALSE, warning = FALSE}

@@

full_data_path <- "data/FullHBForagingData.csv"
data_raw <- tibble(read.csv(full_data_path))

@@

# london area
london <- st_read(
  "shapefiles/London_Ward.shp"
)

@@

ggsave(
  plot = merge_plot,
  filename = "results/figures/site_map.png",
  width = 90,
  height = 110,
  units = "mm",
  dpi = 300
)

@@

ggsave(
  plot = sites_model_plot,
  filename = "results/figures/sites_model_plot.png",
  width = 183,
  height = 190,
  units = "mm",
  dpi = 300
)
```

```{r, other-map-plots, include=FALSE, message = FALSE, warning = FALSE}
# other map plots
inset <- ggplot() +
  geom_sf(
    data = worldmap,
    fill = "grey90",
    color = "#4b4949d0"
  ) +
  geom_sf(
    data = london,
    fill = "#ced0cffc",
    lwd = 0
  ) +
  coord_sf(
    xlim = c(points_area[[1]] - 0.1, points_area[[3]] + 0.1),
    ylim = c(points_area[[2]] - 0.1, points_area[[4]] + 0.1)
  ) +
  geom_point(
    data = map_data,
    aes(x = lon, y = lat, shape = model, colour = model), size = 1.5
  ) +
  annotation_north_arrow(
    location = "tr", which_north = "true",
    style = north_arrow_fancy_orienteering,
    height = unit(10, "mm"),
    width = unit(10, "mm")
  ) +
  annotation_scale(
    location = "br",
    text_cex = 3
  ) +
  scale_shape_manual(values = c(1, 2)) +
  scale_colour_manual(values = c("black", "red")) +
  theme_nothing() +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(x = NULL, y = NULL) +
  theme(
    panel.border = element_rect(color = "black", fill = NA, size = .5),
    text = element_text(size = 42)
  )

@@

ggsave(
  plot = merge_plot,
  filename = "results/figures/site_map.png",
  width = 183,
  height = 190,
  units = "mm",
  dpi = 300
)

@@

ggsave(
  plot = sites_model_plot,
  filename = "results/figures/results_model_plot.png",
  width = 183,
  height = 190,
  units = "mm",
  dpi = 300
)
```

### 2021-10-28 analysis/landuse_effect_prop_scouts.Rmd
started converting other rmwrkdown files into having a nice htlm knitting

```{r get-data, include=FALSE}
# load results of model fit (code to generate found in fit_models.Rmd)
sitedata <- readRDS("results/site_fit_results.Rda")

@@

```{r check_urban}

@@

```{r, peek_urban}
# look at data
print(urban_lu)
```

@@

```{r, urban_pca}
# extract just the lans-use columns
urban.pcadata <- urban_lu[, 3:length(names(urban_lu))]

@@

```{r, urban_pca_scree}
# extract the sd, loadings and scores of the data
sd <- upca$sdev
loadings <- upca$rotation
rownames(loadings) <- colnames(urban.pcadata)
scores <- upca$x

@@

```{r, urban_pca_loadings}
print(loadings[, 1:3])
sqrt(1 / ncol(urban.pcadata)) # cut off for top loadings
```

@@

```{r, urban_p_transform}
n <- length(urban_lu$p)
y <- (urban_lu$p * (n - 1) + 0.5) / n # transform 1 inflated data
mydata <- cbind(y, data.frame(upca$x))
colnames(mydata)[1] <- "transformed.prop.scouts"
```

@@

```{r, urban_pca_regression_3p}
mydata.mod.3 <- betareg(
  transformed.prop.scouts ~ PC1 + PC2 + PC3,
  data = mydata
)
summary(mydata.mod.3)

@@

```{r, urban_pca_regression_p3}
mydata.mod.p3 <- betareg(
  transformed.prop.scouts ~ PC3,
  data = mydata
)
summary(mydata.mod.p3)

@@

```{r, urban_pca_regression_lrt}
mydata.mod.p1_p2 <- betareg(
  transformed.prop.scouts ~ PC2 + PC3,
  data = mydata
)

@@

```{r, urban_pls}
n <- length(urban_lu$p)
y <- (urban_lu$p * (n - 1) + 0.5) / n
urban.pcadata$p_transform <- y

model <- plsRbeta(
  p_transform ~ .,
  data = urban.pcadata,
  nt = ncol(urban.pcadata) - 2,
  model = "pls-beta"
)
print(model)
```

@@

```{r, urban_pls_loadings}
print(model$pp[, 1:3])
sqrt(1 / (ncol(urban.pcadata) - 1))

@@

```{r, urban_pls_p1}

@@

```{r, urban_pls_3p}

@@

```{r, urban_pls_lrt}
plsbetamodel.2 <- betareg(
  prop.scouts ~ Comp_.1 + Comp_.2,
  plsdata
)

@@

```{r, check_rural}

@@

```{r, peek_rural}
print(rural_lu)
```

@@

```{r, rural_pca_scree}
# extract just the lans-use columns
rural.pcadata <- rural_lu[, 3:length(names(rural_lu))]

@@

```{r, rural_pca_loadings}
print(loadings[, 1:4])
sqrt(1 / ncol(rural.pcadata)) # cut off for top loadings
```

@@

```{r, rural_transform_p}
n <- length(rural_lu$p)
y <- (rural_lu$p * (n - 1) + 0.5) / n # transform 1 inflated data
mydata <- cbind(y, data.frame(rpca$x))
colnames(mydata)[1] <- "transformed.prop.scouts"
```

@@

```{r, rural_pca_4p}
mydata.mod.4 <- betareg(
  transformed.prop.scouts ~ PC1 + PC2 + PC3 + PC4,
  data = mydata
)
summary(mydata.mod.4)

@@

```{r, rural_pca_p1}

@@

```{r, rural_pca_p3}
mydata.mod.p3 <- betareg(
  transformed.prop.scouts ~ PC3,
  data = mydata
)
summary(mydata.mod.p3)

@@

```{r, rural_pca_lrt}
lrtest(mydata.mod.p1, mydata.mod.4)
```

@@

```{r, rural_pls}
n <- length(rural_lu$p)
y <- (rural_lu$p * (n - 1) + 0.5) / n
rural.pcadata$p_transform <- y

model <- plsRbeta(
  p_transform ~ .,
  data = rural.pcadata,
  nt = ncol(rural.pcadata) - 1,
  model = "pls-beta"
)
print(model)
```

@@

```{r, rural_pls_loadings}
model$pp[, 1:3]
sqrt(1 / (ncol(rural.pcadata) - 1))
```

@@

```{r, rural_pls_p1}

@@

```{r, rural_pls_4p}

@@

```{r, rural_pls_lrt}
plsbetamodel.2 <- betareg(
  prop.scouts ~ Comp_.1 + Comp_.2,
  plsdata
)

@@

```{r}

@@

model <- plsRbeta(
  p_transform ~ .,
  data = joined.pcdata,
  nt = ncol(joined.pcdata) - 1,
  model = "pls-beta"
)

>>>>

```{r get-data, include=FALSE, message = FALSE, warning = FALSE}
# load results of model fit (code to generate found in fit_models.Rmd)
sitedata <- readRDS("results/site_fit_results.Rda")

@@

```{r check_urban, message = FALSE, warning = FALSE, error = TRUE}

@@

```{r, peek_urban, message = FALSE, warning = FALSE}
# look at data
print(urban_lu)
```

@@

```{r, urban_pca, message = FALSE, warning = FALSE}
# extract just the lans-use columns
urban.pcadata <- urban_lu[, 3:length(names(urban_lu))]

@@

```{r, urban_pca_scree, message = FALSE, warning = FALSE}
# extract the sd, loadings and scores of the data
sd <- upca$sdev
loadings <- upca$rotation
rownames(loadings) <- colnames(urban.pcadata)
scores <- upca$x

@@

```{r, urban_pca_loadings, message = FALSE, warning = FALSE}
print(loadings[, 1:3])
sqrt(1 / ncol(urban.pcadata)) # cut off for top loadings
```

@@

```{r, urban_p_transform, message = FALSE, warning = FALSE}
n <- length(urban_lu$p)
y <- (urban_lu$p * (n - 1) + 0.5) / n # transform 1 inflated data
mydata <- cbind(y, data.frame(upca$x))
colnames(mydata)[1] <- "transformed.prop.scouts"
```

@@

```{r, urban_pca_regression_3p, message = FALSE, warning = FALSE}
mydata.mod.3 <- betareg(
  transformed.prop.scouts ~ PC1 + PC2 + PC3,
  data = mydata
)
summary(mydata.mod.3)

@@

```{r, urban_pca_regression_p3, message = FALSE, warning = FALSE}
mydata.mod.p3 <- betareg(
  transformed.prop.scouts ~ PC3,
  data = mydata
)
summary(mydata.mod.p3)

@@

```{r, urban_pca_regression_lrt, message = FALSE, warning = FALSE}
mydata.mod.p1_p2 <- betareg(
  transformed.prop.scouts ~ PC2 + PC3,
  data = mydata
)

@@

```{r, urban_pls, message = FALSE, warning = FALSE}
n <- length(urban_lu$p)
y <- (urban_lu$p * (n - 1) + 0.5) / n
urban.pcadata$p_transform <- y

model <- plsRbeta(
  p_transform ~ .,
  data = urban.pcadata,
  nt = ncol(urban.pcadata) - 2,
  model = "pls-beta",
  verbose = FALSE
)
print(model)
```

@@

```{r, urban_pls_loadings, message = FALSE, warning = FALSE}
print(model$pp[, 1:3])
sqrt(1 / (ncol(urban.pcadata) - 1))

@@

```{r, urban_pls_p1, message = FALSE, warning = FALSE}

@@

```{r, urban_pls_3p, message = FALSE, warning = FALSE}

@@

```{r, urban_pls_lrt, message = FALSE, warning = FALSE}
plsbetamodel.2 <- betareg(
  prop.scouts ~ Comp_.1 + Comp_.2,
  plsdata
)

@@

```{r, check_rural, message = FALSE, warning = FALSE, error = TRUE}

@@

```{r, peek_rural, message = FALSE, warning = FALSE}
print(rural_lu)
```

@@

```{r, rural_pca_scree, message = FALSE, warning = FALSE}
# extract just the lans-use columns
rural.pcadata <- rural_lu[, 3:length(names(rural_lu))]

@@

```{r, rural_pca_loadings, message = FALSE, warning = FALSE}
print(loadings[, 1:4])
sqrt(1 / ncol(rural.pcadata)) # cut off for top loadings
```

@@

```{r, rural_transform_p, message = FALSE, warning = FALSE}
n <- length(rural_lu$p)
y <- (rural_lu$p * (n - 1) + 0.5) / n # transform 1 inflated data
mydata <- cbind(y, data.frame(rpca$x))
colnames(mydata)[1] <- "transformed.prop.scouts"
```

@@

```{r, rural_pca_4p, message = FALSE, warning = FALSE}
mydata.mod.4 <- betareg(
  transformed.prop.scouts ~ PC1 + PC2 + PC3 + PC4,
  data = mydata
)
summary(mydata.mod.4)

@@

```{r, rural_pca_p1, message = FALSE, warning = FALSE}

@@

```{r, rural_pca_p3, message = FALSE, warning = FALSE}
mydata.mod.p3 <- betareg(
  transformed.prop.scouts ~ PC3,
  data = mydata
)
summary(mydata.mod.p3)

@@

```{r, rural_pca_lrt, message = FALSE, warning = FALSE}
lrtest(mydata.mod.p1, mydata.mod.4)
```

@@

```{r, rural_pls, message = FALSE, warning = FALSE}
n <- length(rural_lu$p)
y <- (rural_lu$p * (n - 1) + 0.5) / n
rural.pcadata$p_transform <- y

model <- plsRbeta(
  p_transform ~ .,
  data = rural.pcadata,
  nt = ncol(rural.pcadata) - 1,
  model = "pls-beta",
  verbose = FALSE
)
print(model)
```

@@

```{r, rural_pls_loadings, message = FALSE, warning = FALSE}
model$pp[, 1:3]
sqrt(1 / (ncol(rural.pcadata) - 1))
```

@@

```{r, rural_pls_p1, message = FALSE, warning = FALSE}

@@

```{r, rural_pls_4p, message = FALSE, warning = FALSE}

@@

```{r, rural_pls_lrt, message = FALSE, warning = FALSE}
plsbetamodel.2 <- betareg(
  prop.scouts ~ Comp_.1 + Comp_.2,
  plsdata
)

@@

```{r, combined-pls, message = FALSE, warning = FALSE}

@@

model <- plsRbeta(
  p_transform ~ .,
  data = joined.pcdata,
  nt = ncol(joined.pcdata) - 1,
  model = "pls-beta",
  verbose = FALSE
)
